Handbook streamFind: backend

### 2022-07-20 handbooks/Handbook_backend_streamFind.Rmd
Large framework update 07

```{r libraries, include=FALSE, cache=FALSE}

library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(streamFind)

#devtools::load_all("../")
```

@@

The streamFind package is a backend (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing. The major focus of streamFind is data processing for environmental and quality studies of the water cycle. The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data. The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities (modules). This handbook focuses on describing and demonstrating the backend of the streamFind package.

@@

The streamFind package uses S4 class objects. The S4 classes, main methods and usage are presented and exemplified in this section. An overview of the S4 classes and their hierarchy is presented in Figure \@ref(fig:classes-overview) using MS analyses as an example. Other data types, such as from UV and Raman analyses, would lead to different classes (e.g., _uvData_ and _uvAnalysis_ and _ramanData_ and _ramanAnalysis_, respectively) but the hierarchy is similar as shown for MS data.

```{r classes-overview, echo=FALSE, fig.cap="Overview of the S4 classes and their hierarchy with emphasis on MS analyses.", out.width="120%"}
knitr::include_graphics(paste0(getwd(), "/hb_backend_figures/classes_overview.png"))
```

## Basic terminalogy

In streamFind, the terms listed below are consistently applied throughout the objects and functionalities.

* __project__ is the project title;
* __analysis__ is the name or the actual analysis file;
* __replicate__ is the name of an analysis replicate group;
* __blank__ is the name of a blank analysis replicate group ot be used for blank (blind) subtration;
* __trace__ correspond to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;  
* __spectra__ in MS is a table with a spectrum for each cycle time (i.e., retention time);
* __EIC__ is an extracted ion chromatograms (retention time vs intensity);
* __TIC__ is a total ion chromatograms (retention time vs intensity);
* __XIC__ is a three dimensional ion chromatogram (_m/z_ vs retention time vs intensity);
* __peak__ correspond to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file;
* __feature__ correspond to a group of corresponding peaks across analysis files;
* __MS2__ correspond to second order fragmentation data as acquired in tandem MS;
(...)

@@

The project creation is demonstrated below for three HRMS files (triplicate) where a set of substances were spiked. The project path can be assign as argument but left as default (i.e., the working directory) for this example.
The initial S4 class created in the `newStreamProject` function is _streamProject_.

```{r newStreamProject-call, message=FALSE, results='hide'}
exS1 <- newStreamProject(files = files[1:3])
```

@@

Because the added files were all with mzML format, the sub-class is directly defined as _msData_, adding additional structure for further processing of MS data. Other data formats/structure will lead to different sub-classes (e.g., _uvData_ and _ramanData_) but is not yet implemented within streamFind.

```{r msData_streamProject_classes}
is(exS1)
```

The _msData_ structure includes the _streamProject_ class plus the `features` slot. The project, date and path slots correspond to the project title, creation date and directory in disk, respectively. The analyses slot is a list with a _msAnalysis_ S4 class object per MS file added. The structure of the _msAnalysis_ class is described in the next sub-chapter.

The `features` `r # TODO make msFeatures after peaks extraction and storage`` 
which includes the _msFeatures_ S4 class(presented in section X).

@@

```{r structure_msData}
str(exS1, max.level = 2)
```

@@

```{r msAnalysis_structure}
exA1 <- getAnalyses(exS1, 1)
str(exA1, max.level = 2)
```

@@

```{r metadata_structure}
t(streamFind::metadata(exS1, analyses = 1))

#applied directly to the msAnalysis S4 class
t(streamFind::metadata(exA1))
```

@@

# Low level data access and plotting

The tools for low level access and plotting (i.e., traces and chromatograms) share a target search approach. The data search is based on pre-defined time and mass (inc. _m/z_) targets. For traces (i.e., raw MS mass traces), time (in seconds) and _m/z_ (in Da) ranges are given for accessing and plotting. Access to mass traces is relevant for further development of data processing steps/functions and for visualization of raw data, easing the evaluation of processing steps (e.g., peaks wrongly not found during peak picking).

@@

As aforementioned, access to raw data is based on defining time and _m/z_ targets. The function `makeTargets` is used for comprehensively build targets. As shown below, diverse ways can be used to assemble targets, similarly returning a uniform `data.table`. Accessing and plotting S4 methods in streamFind use the `makeTargets` function to collect data. Therefore, the arguments `mz`, `rt`, `ppm`, `sec` and `id` are present in most S4 methods for accessing and plotting data as demonstrated in the following sections.

```{r make_targets}
#case 1
mz01 <- c(247.1651, 239.0628)
rt01 <- c(839, 937)
id01 <- c("target1", "target2")
targets01 <- makeTargets(mz = mz01, rt = rt01, ppm = 20, sec = 60, id = id01)

targets01

#case 2
mz02 <- data.frame(mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652))
rt02 <- data.frame(rtmin = c(809, 907), rtmax = c(869, 967))
targets02 <- makeTargets(mz = mz02, rt = rt02)

targets02

#case 3
mz03 <- data.frame(
  #id = c("target1", "target2"),
  mz = c(247.1651, 239.0628),
  rt = c(839, 937)
)
targets03 <- makeTargets(mz = mz03, ppm = 20, sec = 60)

targets03

#case 4
mz04 <- data.frame(
  id = c("target1", "target2"),
  mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652),
  rtmin = c(809, 907), rtmax = c(869, 967)
)
targets04 <- makeTargets(mz = mz04)

targets04
```

@@

```{r EICs_method}
#EIC of targets from first analysis using msAnalysis class
eic01 <- EICs(exA1, mz = targets04)
head(eic01, 6)

#EIC of targets from second analyses using msData class
eic02 <- EICs(exS1, analyses = 2, mz = targets04)
head(eic02, 6)
```

@@

```{r plotEICs_static}
#plot the produced EIC data.table
plotEICs(eic01)
```

```{r plotEICs_interactive}
#using the msData class object and targets but with interactive plot
plotEICs(exS1, analyses = NULL, mz = targets04, colorBy = "analyses", interactive = TRUE)

#the same plot as the first for data.table but interactive
plotEICs(exA1, mz = targets04, interactive = TRUE)
```

@@

```{r TICs_method}
# TIC for both analyses in the msData class object
tic01 <- TICs(exS1)
tic01

#TIC of the msAnalysis object
head(TICs(exA1), 6)
```

@@

```{r plotTICs_method}
#using a data.table as obatined by TICs
plotTICs(tic01, interactive = TRUE)

#using the msData object
plotTICs(exS1, colorBy = "replicates")

#using the msAnalysis
plotTICs(exA1)
```

@@

```{r XICs_method}
#using the msData class object
xic01 <- XICs(exS1, mz = targets04)
xic01

#with a msAnalysis class object
xic02 <- XICs(exA1, mz = mz01, rt = rt01, ppm = 20, sec = 60)
head(xic02, 3)
```

@@

```{r plotXICs_method}
#using the output of XICs S4 method (a data.table)
plotXICs(xic01, legendNames = c("target number 1", "target number 2"), plotTargetMark = TRUE,
 targetsMark = targets04[, c("mz", "rt")], ppmMark = 5, secMark = 10, numberRows = 2)

#using the msData class and plotting the first target from the first analysis
plotXICs(exS1[1], mz = targets04[1, ])

#for the msAnalysis, the second target with wide m/z and rt ranges
plotXICs(exA1, mz = mz01[2], rt = rt01[2], ppm = 500, sec = 120)
```

@@

```{r loadRawData}
#load raw data to an msAnalysis object
exA1 <- loadRawData(exA1)

#example of accessing the six most intense fragments of the first target with msAnalysis
s_data <- spectra(exA1)
s_data <- s_data[preMZ >= targets04$mzmin[1] & preMZ <= targets04$mzmax[1], ]
s_data <- s_data[rt >= targets04$rtmin[1] & rt <= targets04$rtmax[1], ]
s_data <- s_data[order(intensity, decreasing = TRUE), ]
s_data[1:6, ]

#load raw data to an msData object.
exS1 <- loadRawData(exS1, analyses = NULL)

#example of accessing the two most intense fragments of the first target with msData
s_data_2 <- spectra(exS1)
s_data_2 <- s_data_2[preMZ >= targets04$mzmin[1] & preMZ <= targets04$mzmax[1], ]
s_data_2 <- s_data_2[rt >= targets04$rtmin[1] & rt <= targets04$rtmax[1], ]
s_data_2 <- s_data_2[order(intensity, decreasing = TRUE), ]
s_data_2[1:4, ]
```

@@

The basic workflow (Figure \@ref(fig:workflow-scheme)) includes the following steps: (1) project creation, (2) data conversion (when required/applicable), (3) assignment of analysis replicate names and respective blank analysis replicates, (4) peak finding (or peak picking), (5) peak alignment and grouping into features across analyses, (6) filling of peaks missing in features, (7) annotation of isotopic and adduct features and (8) feature quality evaluation and (9) filtering. The usage of the S4 class objects within the basic workflow (pre-processing) for HRMS data is presented in this section with examples.

@@

```{r inspecting_peaks}
#access to peaks in msData based on built targets as used for extracting EICs
peaks(exS2_pp[c(1, 4)], mz = mz01[1], rt = rt01[1], ppm = 10, sec = 30)

#access to peak in msAnalysis
peaks(exA2, mz = targets04)

#plotting peaks in msData
plotPeaks(exS2_pp, mz = targets04[2, ], interactive = TRUE, colorBy = "analyses")

#plotting peaks from msAnalysis
plotPeaks(exA2, mz = targets04, interactive = FALSE)

#map plot for peak time and m/z dimensions for msAnalysis
mapPeaks(exA2, mz = targets04[1, ], xlim = 30, ylim = 0.001)

##map plot for peak time and m/z dimensions for msData
mapPeaks(exS2_pp, mz = targets04[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
```

@@

```{r inspecting_features}
#getter for feature intensities and intensity deviations
features(exS2_pg, mz = targets04)

#getter for all the feature metadata by setting complete to TRUE
t(features(exS2_pg_2, mz = targets04[1, ], complete = TRUE))

#getter for features with intensities for each sample by setting average to FALSE
features(exS2_pg_2, mz = targets04[1, ], average = FALSE)

#plotting features
plotFeatures(exS2_pg, mz = targets04, colorBy = "targets", interactive = FALSE)

#plotting features for replicates and interactive
plotFeatures(exS2_pg, mz = targets04, colorBy = "replicates", interactive = TRUE)

#plot the individual peaks in features and for each analysis shows the time deviation.
plotFeaturePeaks(exS2_pg, mz = targets04)
```

@@

```{r extract_msn}
#head of MS2 data for target 1 in a msAnalysis object
head(MS2s(exA1, mz = targets04[1, ], settings = settings_msn))

#ploting the MS2 data of the first target in a msAnalysis object
plotMS2s(exA1, mz = targets04[1, ], settings = settings_msn, interactive = FALSE)

#most intense MS2 data traces for both targets in a msData object
s_data_3 <- MS2s(exS1, mz = targets04, settings = settings_msn)
s_data_3 <- s_data_3[order(intensity, decreasing = TRUE), ]
head(s_data_3)

#ploting the MS2 data of the first target in a msAnalysis object
plotMS2s(exS1, mz = targets04, settings = settings_msn, interactive = TRUE, colorBy = "targets")
```

@@

# trimSpectraFilesMZR(files = choose.files(), rtr = c(500, 1000), mzr = c(200, 800), onlyMS1 = TRUE)
# 
# mz_05 <- data.frame(
#   retmin = c(809, 907), retmax = c(869, 967),
#   mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652)
# )
# 
# #test <- patRoon::getEICs(msa_1@file, mz_05)
# 
# test <- extractEICs(object, mz = targets04)
# test_xic <- extractXICs(object, mz = targets04)

@@

# Instructions

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
Diverse cheatsheets can be found in <https://www.rstudio.com/resources/cheatsheets/>.
This section gives quick instructions to display text, code, tables and figures.

## Text

Text in R markdown has several inline codes for costumization.
The book in <https://bookdown.org/yihui/rmarkdown/> offers a detailed guideline.
Moreover, a cheatsheet can be downdloaded with the following link <https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf>.

## Code

Code is automaticelly embedded unless `echo` chunk option is set to `FALSE`.
Evaluation of code in a given chunk can be skiped by setting `eval` to `FALSE`.
Code chunks should always be unically named.

```{r exCode, echo=FALSE, eval=FALSE, include=FALSE}

#Example of code chunk

list(a = 1:6, b = 1:10)

<br>

## Tables

The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exTable). Consult <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html> for other costumization options.
Table caption is added with the argument `caption` of the `kable` function.

```{r exTable, echo=TRUE, eval=TRUE, include=TRUE}

df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)

knitr::kable(df, caption = "Example of table caption.") %>%
kable_styling(
  font_size = 12,
  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
  fixed_thead = TRUE
)

<br>

## Figures

Figures/plots can also by embedded using diverse packages.
The packages `ggplot2` and `plotly` (interactive) are recommended as they are versatile and easy to use.
Cheatsheet for `ggplot2` can be found in <https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf>.
For `plotly` the website <https://plotly.com/r/> offers examples and a cheatsheet can be downdloaded in <https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf>.
Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the R code that generate the plots.

### Example `ggplot2`

```{r exGgplot, fig.cap="Plot caption example"}

ggplot(pressure, aes(temperature, pressure)) +
geom_point() +
theme_bw()

### Example `plotly`

```{r exPlotly, fig.cap="Plot caption example"}

plot_ly(pressure, x = ~temperature, y = ~pressure, type = "scatter", mode = "markers+lines")

<br> <br>

# Signature

The report was produced on `r Sys.time()`. The data and report are stored in of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.

<br>

@@

<style>
body, p {
  color: black;
  font-family: Arial;
  text-align: justify;
  font-size: 14pt;
}
h1{
  font-size: 26pt;
  font-style: bold;
}
h2{
  font-size:20pt;
  font-style: bold;
}
h3{
  font-size:16pt;
  font-style: bold;
}
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #239B56;
}
caption {
  color: black;
  font-size: 1.0em;
}
</style>

>>>>

```{r libraries, include=FALSE, cache=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(streamFind)
```

@@

The streamFind package is a backend (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing. The major focus of streamFind is data processing for environmental and quality studies. The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data. The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities (modules). This handbook focuses on describing and demonstrating the backend of the streamFind package.

@@

The streamFind package uses S4 class objects (e.g., _streamProject_, _msData_, _msAnalysis_). The S4 classes and their main getter and setter methods are presented and exemplified in this section. An overview of the S4 classes and their hierarchy is presented in Figure \@ref(fig:classes-overview) using MS analyses as an example. Other data types, such as from UV and Raman analyses, would lead to different classes (e.g., _uvData_ and _uvAnalysis_ and _ramanData_ and _ramanAnalysis_, respectively) but the hierarchy is similar as shown for MS data.

```{r classes-overview, echo=FALSE, fig.cap="Overview of the S4 classes and their hierarchy with emphasis on MS analyses.", out.width="95%"}
knitr::include_graphics(paste0(getwd(), "/hb_backend_figures/classes_overview.png"))
```

## Nomenclature

The nomenclature listed below is consistently applied throughout the objects and arguments of functionalities in the streamFind package. Other less relevant terms are used but not listed. For further information please consult the specific documentation of each object or functionality when required using the `?x` method, where x is the name of the S4 class, S4 method or function.

* _streamProject_ refers to the broader S4 class applied in the streamFind package and holds basic project information, see \@ref(streamproject-class);
* _msData_ refers to the S4 subclass of _streamProject_ when MS analyses (i.e., MS files) are added to the project, see \@ref(msdata-class);
* _msAnalysis_ refers to the S4 class of the object generated for each MS file added to the _msData_ object in the slot "analyses";
* `projectInfo` is the method to obtain the basic information (i.e., title, date and path) of a _streamProject_ object;
* `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the _streamProject_;
* `replicate` is the name of an analysis replicate group and the method to get the replicate names;
* `blank` is the name of a blank analysis replicate group to be used for blank (blind) subtraction and the method to get blank replicate names;
* __traces__ corresponds to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;  
* `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time
`r # TODO add chromatograms entry and make method`
* __EIC__ is an extracted ion chromatogram (retention time vs intensity) and can be obtained with the method `EICs`;
* __TIC__ is a total ion chromatograms (retention time vs intensity) and can be obtained with the method `TICs`;
* __XIC__ is a three dimensional ion chromatogram (_m/z_ vs retention time vs intensity) and can be obtained with the method `XICs`;
* __peak__ corresponds to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file and can be accessed with the method `peaks`;
* __feature__ corresponds to a group of corresponding peaks across analysis files/analyses and can be accessed with the method `features`;
* __MS2__ corresponds to second order fragmentation data as acquired in tandem MS or MS/MS mode and can be accessed with the method `MS2s`.

(...)

@@

The project creation is demonstrated below for the three MS files (triplicate) where a set of chemicals were spiked (see \@ref(spiked-chemicals)). The project path can be assign as argument but for the example below it was left as default (i.e., the working directory).
The initial S4 class created in the `newStreamProject` function is _streamProject_. The subclass is then defined according to the type of files added. Note that multiple file types are not possible as a subclass cannot be defined.

```{r newStreamProject-call, message=FALSE, results='hide'}
sp1 <- newStreamProject(files = files[1:3], title = "Example_01")
```

@@

Because the added files were all with mzML format (i.e., MS files), the subclass is directly defined as _msData_. The structure for further processing data is then set for MS data. Other data formats/structure will lead to different subclasses (e.g., _uvData_ and _ramanData_) but is not yet implemented within the streamFind package.

```{r msData_streamProject_classes}
is(sp1)
```

The _msData_ structure includes the _streamProject_ class plus the `features` slot. The project, date and path slots correspond to the project title, creation date and directory in disk, respectively. The analyses slot is a list with a _msAnalysis_ S4 class object per MS file added. The structure of the _msAnalysis_ class is described in the next sub-chapter. The `features` slot is a _msFeatures_ S4 class object to hold the results from alignment, grouping and annotation of peaks across analyses.

@@

```{r structure_msData}
str(sp1, max.level = 2)
```

@@

```{r msAnalysis_structure}
a1 <- getAnalyses(sp1, 1)
str(a1, max.level = 2)
```

@@

```{r metadata_structure}
t(streamFind::metadata(sp1, analyses = 1))

#applied directly to the msAnalysis S4 class
t(streamFind::metadata(a1))
```

@@

# Raw data access

The tools for raw data access and plotting (i.e., spectra and chromatograms) share a target search approach. The data search is based on pre-defined time and mass (inc. _m/z_) targets. For traces (i.e., raw MS mass traces), time (in seconds) and _m/z_ (in Da) ranges are given for accessing and plotting. Access to mass traces is relevant for further development of data processing steps/functions and for visualization of raw data, easing the evaluation of processing steps (e.g., peaks wrongly not found during peak picking).

@@

As aforementioned, access to raw data is based on defining _m/z_ and retention time targets. The function `makeTargets` is used for comprehensively build targets. As shown below, multiple ways can be used to assemble targets but similarly returning a uniform `data.table` class object with a target list. Accessing and plotting S4 methods in streamFind use the `makeTargets` function to collect raw data from the analysis files. Therefore, the arguments `mz`, `rt`, `ppm`, `sec` and `id` are present in most S4 methods for accessing and plotting data as demonstrated in the following sections.

```{r make_targets}
#case 1
mz1 <- c(247.1651, 239.0628)
rt1 <- c(839, 937)
id1 <- c("target1", "target2")
targets1 <- makeTargets(mz = mz1, rt = rt1, ppm = 20, sec = 60, id = id1)

targets1

#case 2
mz2 <- data.frame(mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652))
rt2 <- data.frame(rtmin = c(809, 907), rtmax = c(869, 967))
targets2 <- makeTargets(mz = mz2, rt = rt2)

targets2

#case 3
mz3 <- data.frame(
  id = c("target1", "target2"),
  mz = c(247.1651, 239.0628),
  rt = c(839, 937)
)
targets3 <- makeTargets(mz = mz3, ppm = 20, sec = 60)

targets3

#case 4
mz4 <- data.frame(
  id = c("target1", "target2"),
  mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652),
  rtmin = c(809, 907), rtmax = c(869, 967)
)
targets4 <- makeTargets(mz = mz4)

targets4
```

@@

```{r EICs_method}
#EIC of targets from first analysis using msAnalysis class
eic1 <- EICs(a1, mz = targets4)
head(eic1, 6)

#EIC of targets from second analyses using msData class
eic2 <- EICs(sp1, analyses = 2, mz = targets4)
head(eic2, 6)
```

@@

```{r plotEICs_static}
#plot the produced EIC data.table
plotEICs(eic1)
```

```{r plotEICs_interactive}
#using the msData class object and targets but with interactive plot
plotEICs(sp1, analyses = NULL, mz = targets4, colorBy = "analyses", interactive = TRUE)

#the same plot as the first for data.table but interactive
plotEICs(a1, mz = targets4, interactive = TRUE)
```

@@

```{r TICs_method}
# TIC for both analyses in the msData class object
tic1 <- TICs(sp1)
tic1

#TIC of the msAnalysis object
head(TICs(a1), 6)
```

@@

```{r plotTICs_method}
#using a data.table as obatined by TICs
plotTICs(tic1, interactive = TRUE)

#using the msData object
plotTICs(sp1, colorBy = "replicates")

#using the msAnalysis
plotTICs(a1)
```

@@

```{r XICs_method}
#using the msData class object
xic1 <- XICs(sp1, mz = targets4)
xic1

#with a msAnalysis class object
xic2 <- XICs(a1, mz = mz1, rt = rt1, ppm = 20, sec = 60)
head(xic2, 3)
```

@@

```{r plotXICs_method}
#using the output of XICs S4 method (a data.table)
plotXICs(xic1, legendNames = c("target number 1", "target number 2"), plotTargetMark = TRUE,
 targetsMark = targets4[, c("mz", "rt")], ppmMark = 5, secMark = 10, numberRows = 2)

#using the msData class and plotting the first target from the first analysis
plotXICs(sp1[1], mz = targets4[1, ])

#for the msAnalysis, the second target with wide m/z and rt ranges
plotXICs(a1, mz = mz1[2], rt = rt1[2], ppm = 500, sec = 120)
```

@@

```{r loadRawData}
#load raw data to an msAnalysis object
a1 <- loadRawData(a1)

#example of accessing the six most intense fragments of the first target with msAnalysis
s_data <- spectra(a1)
s_data <- s_data[preMZ >= targets4$mzmin[1] & preMZ <= targets4$mzmax[1], ]
s_data <- s_data[rt >= targets4$rtmin[1] & rt <= targets4$rtmax[1], ]
s_data <- s_data[order(intensity, decreasing = TRUE), ]
s_data[1:6, ]

#load raw data to an msData object.
sp1 <- loadRawData(sp1, analyses = NULL)

#example of accessing the two most intense fragments of the first target with msData
s_data_2 <- spectra(sp1)
s_data_2 <- s_data_2[preMZ >= targets4$mzmin[1] & preMZ <= targets4$mzmax[1], ]
s_data_2 <- s_data_2[rt >= targets4$rtmin[1] & rt <= targets4$rtmax[1], ]
s_data_2 <- s_data_2[order(intensity, decreasing = TRUE), ]
s_data_2[1:4, ]
```

@@

The basic workflow (Figure \@ref(fig:workflow-scheme) left column) includes the following steps: (1) project creation, (2) data conversion (when required/applicable), (3) assignment of analysis replicate names and respective blank analysis replicates, (4) peak finding (or peak picking), (5) peak alignment and grouping into features across analyses, (6) filling of peaks missing in features, (7) annotation of isotopic and adduct features and (8) feature quality evaluation and (9) filtering. The usage of the S4 class objects within the basic workflow (pre-processing) for HRMS data is presented in this section with examples. The right column in Figure \@ref(fig:workflow-scheme illustrates the implementation of modular functionalities for assembly of data processing workflows. The backend for modular assembly and framework of workflows in streamFind is described and demonstrated in section \@ref(workflows).

@@

```{r inspecting_peaks}
#access to peaks in msData based on built targets as used for extracting EICs
peaks(exS2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)

#access to peak in msAnalysis
peaks(exA2, mz = targets4)

#plotting peaks in msData
plotPeaks(exS2_pp, mz = targets4[2, ], interactive = TRUE, colorBy = "analyses")

#plotting peaks from msAnalysis
plotPeaks(exA2, mz = targets4, interactive = FALSE)

#map plot for peak time and m/z dimensions for msAnalysis
mapPeaks(exA2, mz = targets4[1, ], xlim = 30, ylim = 0.001)

##map plot for peak time and m/z dimensions for msData
mapPeaks(exS2_pp, mz = targets4[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
```

@@

```{r inspecting_features}
#getter for feature intensities and intensity deviations
features(exS2_pg, mz = targets4)

#getter for all the feature metadata by setting complete to TRUE
t(features(exS2_pg_2, mz = targets4[1, ], complete = TRUE))

#getter for features with intensities for each sample by setting average to FALSE
features(exS2_pg_2, mz = targets4[1, ], average = FALSE)

#plotting features
plotFeatures(exS2_pg, mz = targets4, colorBy = "targets", interactive = FALSE)

#plotting features for replicates and interactive
plotFeatures(exS2_pg, mz = targets4, colorBy = "replicates", interactive = TRUE)

#plot the individual peaks in features and for each analysis shows the time deviation.
plotFeaturePeaks(exS2_pg, mz = targets4)
```

@@

```{r extract_msn}
#head of MS2 data for target 1 in a msAnalysis object
head(MS2s(a1, mz = targets4[1, ], settings = settings_msn))

#ploting the MS2 data of the first target in a msAnalysis object
plotMS2s(a1, mz = targets4[1, ], settings = settings_msn, interactive = FALSE)

#most intense MS2 data traces for both targets in a msData object
s_data_3 <- MS2s(sp1, mz = targets4, settings = settings_msn)
s_data_3 <- s_data_3[order(intensity, decreasing = TRUE), ]
head(s_data_3)

#ploting the MS2 data of the first target in a msAnalysis object
plotMS2s(sp1, mz = targets4, settings = settings_msn, interactive = TRUE, colorBy = "targets")
```

@@

# trimSpectraFilesMZR(files = choose.files(), rtr = c(500, 1000), mzr = c(200, 800), onlyMS1 = TRUE)
# 
# mz_05 <- data.frame(
#   retmin = c(809, 907), retmax = c(869, 967),
#   mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652)
# )
# 
# #test <- patRoon::getEICs(msa_1@file, mz_05)
# 
# test <- extractEICs(object, mz = targets4)
# test_xic <- extractXICs(object, mz = targets4)

@@

@@
